---
title: "Data Mining Homework 1"
author: "Arman Zakaryan"
date: "September 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1


```{r cars}
load(file = 'Cars93.rda')
dim(df)
```

## Problem 2


```{r }
head(df, n=5)
```

## Problem 3


```{r }
df[which.max(df$Max.Price),c(2,3)]
df[which.min(df$Min.Price),c(2,3)]
```

## Problem 4

The class choice is not ideal, so we can change it to factor with the following layers:

```{r }
class(df$Origin)
df$Origin <- factor(df$Origin,labels=c('non-USA','USA'))
head(df)
class(df$Origin)
```

## Problem 5


```{r }
class(df$Type)
df$Type
```
So the class is $factor$ and the variable is ordinal, because it indicates the size of given car (sort of a quantitative measure).


## Problem 6

Here we observe a positive relationship
```{r }
set.seed(1)
df_small <- df[sample(nrow(df),50),]
plot(x=df_small$EngineSize, y=df_small$Price, 
     main='Scatterplot of Engine Size and Price',
     xlab='Engine Size', ylab='Price')
```

## Problem 7

Here we see a negative relationship
```{r }
plot(x=df_small$MPG.city, y=df_small$Price, 
     main='Scatterplot of MPG in City and Price',
     xlab='MPG in City', ylab='Price')
```

## Problem 8


```{r }
plot(x=df_small$Origin, 
     main='Bar Graph of Origin',
     xlab='Origin')
```

Slightly more non-USA cars.


## Problem 9

The most appropriate graph to represent the distribution is histogram. It shows the frequency of each range of values in the dataset.
The higher "bars" rise, the higher is frequency.

```{r }
hist(df$Price, main='Histogram of Prices', breaks=10,
     xlab='Price', ylab='Number of Cars')
```

## Problem 10

Boxplot is a good graph to explore the outliers. Any data point past whiskers will be considered an outlier.

```{r }
boxpltres <- boxplot(df$Price, main='Boxplot of Price', ylab='Price')
length(boxpltres$out)
```

## Problem 11


```{r }
df <- df[df$Price>=boxpltres$stats[1] & df$Price<=boxpltres$stats[5],]
dim(df)


boxpltres2 <- boxplot(df$Max.Price, main='Boxplot of Max Price', ylab='Price')
df <- df[df$Max.Price>=boxpltres2$stats[1] & df$Max.Price<=boxpltres2$stats[5],]
dim(df)
```

5 observations have been omitted, as they were outliers by either Price or Max.Price or both.


## Problem 12


```{r }
aggregate(df$Luggage.room~df$Manufacturer, data = df, mean)
```


## Problem 13


```{r }
aggregate(df$Max.Price~df$Manufacturer, data=df, max)
```


## Problem 14


```{r }
library(ggplot2)
mymidwest <- midwest
p <- ggplot(mymidwest,aes(x=mymidwest$poptotal, 
                     y=mymidwest$area, color=mymidwest$state))+
  geom_point()+
  labs(x='Total Population', y='Area', 
       title = 'Scatterplot of Total Population and Area', 
       color='Color by States')

p
```


## Problem 15


```{r }
p+facet_grid(mymidwest$state~.)
```


## Problem 16

Interesting finds:
1. Percentage of College Educated People is positively correlated with Perentage of people with Profession.
2. Percentage of Black people is positively correlated with percentage of people living below poverty.


```{r }

ggplot(mymidwest,aes(x=mymidwest$percprof, 
                     y=mymidwest$percollege))+
  geom_point()+geom_smooth(method = 'lm', na.rm=TRUE)+
  labs(x='Percent Profession', y='Percentage of C.E.P', 
       title = 'Scatterplot of Percentages of CEP and Professions')

ggplot(mymidwest,aes(x=mymidwest$percblack,
                     y=mymidwest$percbelowpoverty))+
  geom_point()+geom_smooth(method = 'lm', na.rm=TRUE)+
  labs(x='Percentage of Black People', y='Percentage of People Below Poverty', 
       title = 'Scatterplot of Percentages of Black People and People Below Poverty')
  

```


The choice of both is intition and a little exploration of the Correlation Matrix.


## Problem 17

The graph below looks fuzzy, with a hint of negative relationship.
Intuition would make me expect not-so-strong correlation, as in my opinion college provides enough background not to end up in poverty, but arguably not sufficient enough. 

```{r }
ggplot(mymidwest,aes(x=mymidwest$percollege, 
                     y=mymidwest$percpovertyknown))+
  geom_point()+
  labs(x='Percentage of C.E.P', y='Known Percentage of Poverty', 
       title = 'Scatterplot of Percentages of CEP and Poverty')

```


## Problem 18


```{r }
cor(mymidwest$percollege, mymidwest$percpovertyknown)

```

Yes!


## Problem 19

Signs show the nature of relationships between one another. Positive is the variables moving in one direction, negative means the opposite.

```{r }
num_midwest <- mymidwest[,4:13]
cor(num_midwest)
```

The most correlated ones are $poptotal$ and $popwhite$ (DUH).

## Problem 20


```{r }
library(corrplot)
midwestcor <- mymidwest[,c(-2,-3,-ncol(mymidwest))]
corrplot(cor(midwestcor), method = 'shade')
```
